\input{myassignmentpreamble}
\input{yliow}
\input{ciss451}

\renewcommand\TITLE{Final exam (takehome)}
\newcommand\LISTMARGIN{1in}
\renewcommand{\thethm}{\arabic{thm}}


\begin{document}
\topmatter

\newpage
\textsc{Typesetting aligned equations with comments/justifications}

Here's an example of typesetting aligned computations (with justifications).
Suppose I want to prove
$(x + y \cdot z) + (-y) \cdot z) = x$.
And I can only use the following:

Let $(R, +, \cdot, 0, 1)$ be a ring. 
\begin{enumerate}[nosep]
 \li The definition of $R$ being a ring, i.e., the ring axioms of $R$.
 \li Fact 1: If $x \in R$, then $0 \cdot x = 0$.
 \li Fact 2: If $x \in R$, then $0 \cdot x = 0 = x \cdot 0$.
 \li Fact 3: If $x + y = 0$, then $y = -x$.
 \li Fact 4: $y + x = 0$, then $y = -x$.
 \li Fact 5: If $x \in R$, then $-(-x) = x$.
\end{enumerate}

The I will show $(x + y \cdot z) + (-y) \cdot z) = x$
like this:

\begin{align*}
(x + y\cdot z) + (-y) \cdot z &= x + (y \cdot z + (-y) \cdot z) & & \text{by the associativity axiom of $+$} \\
                              &= x + (y + (-y)) \cdot z         & & \text{by the distributivity axiom} \\
                              &= x + 0 \cdot z                  & & \text{by the inverse axiom of $+$} \\
                              &= x + 0                          & & \text{by Fact 2} \\
                              &= x                              & & \text{by neutrality axiom of $+$} \\
\end{align*}

Take a look at the \LaTeX\ code.
The \verb!&! are alignment characters.


%===============================================================================
\newpage
Q1.
What is the ones digit of the following number
\[
1357^{{2468}^{{3579}^{{4680^{5791^{{6802^{7913^{8024^{9135}}}}}}}}}}
\]
A complete proof is required.

\textsc{Solution}

\input{q2.tex}

%===============================================================================
\newpage
Q2.
In this question, you will prove several basic facts about groups.

In the proofs below, you assume use the following:
Let $(G, *, e)$ be a ring. 
\begin{enumerate}[nosep]
 \li The definition of $G$ being a group, i.e., the group axioms of $R$.
 \li Fact 1: Identity element is unique. In other words let $e,e' \in G$
 such that
 \begin{align*} 
  e * x &= x = x * e \\
  e' * x &= x = x * e'   
 \end{align*}  
 for all $x \in G$. Then $e = e'$.
 (This is proposition 202.2.1 in the notes.)
 \li Fact 2: Inverse of an element is unique.
 In other words let $x \in G$. Suppose $y,y' \in G$ such that
 \begin{align*} 
  x * y &= e = y * x \\
  x * y' &= e = y' * x   
 \end{align*}
 Then $y = y'$.
 (This is proposition 202.2.2 in the notes.)
 \li Fact 3: Left cancellation holds.
 In other words,
 let $a,x,y \in G$ such that
 \[
 a * x = a * y
 \]
 then
 \[
 x = y
 \]
 Likewise, if
 \[
 x * a = y * a
 \]
 then
 \[
 x = y
 \]
\end{enumerate}
Do not use any justification other than the axioms and Facts 1-3.

Prove the following
\begin{enumerate}
\item[(a)] $(x^{-1})^{-1} = x$.
\item[(b)] $(x * y)^{-1} = y^{-1} * x^{-1}$.
\end{enumerate}

\textsc{Solution}

\input{q2.tex}

%===============================================================================
\newpage
Q3.
Assume the given facts about groups from Q2.
Furthermore define $x^n$ for $n \geq 0$ as follows:
\[
x^n =
\begin{cases}
e           & \text{ if $n = 0$ } \\
x           & \text{ if $n = 1$ } \\
x^{n-1} * x  & \text{ if $n > 1$ }
\end{cases}
\]
This is from the notes which also contains the definition of $x$ raised to a
negative power.

Prove that
\[
(x^n)^{-1} = (x^{-1})^n
\]
for $n \geq 0$ by induction.
(This above is also true when $n$ is negative. But you need to prove
the above for negative $n$.)


%==============================================================================
\newpage
Q4. Consider the ring $R = (\Z/2)[X]/n$ where $n = X^2 + 1$.
\begin{enumerate}[nosep]

 \item
 What is $|R|$, the size of $R$?

 \item
 Factorize $n = X^2 + 1$ in $(\Z/2)[X]$.

 \item
 For each element $x$ of $R$, write down the multiplicative inverse of $x$.
 Is $R$ a field?
 
\end{enumerate}

\textsc{Solution}

\input{q2.tex}

%==============================================================================
\newpage
Q5.
Can RSA be extended to three primes?
In other words let $p,q,r$ be \textit{three} (not two) primes and let $N = pqr$.
$\phi(N)$ is the Euler totient of $N$.
Let $e,d$ are integers such that $ed \equiv 1 \pmod{N}$.
Let $x$ be an integer.
Then
\[
\left( x^e \right)^d \equiv x \pmod{N} \tag{*}
\]
If the above is not true, provide a counter-example.
Otherwise prove $(*)$.

\textsc{Solution}

\input{q5.tex}


%==============================================================================
\newpage
Q6.
Let $p$ be a prime.
Prove that $\sqrt{p}$ is irrational (i.e., not a fraction) using
the well-ordering principle.
Note: You must use the well-ordering principle.

(Hint: $\sqrt{2}$ is irrational is proven in discrete 1.
That was usually proved using proof by contradiction.
Redesign the proof to use WOP.
Then generalize the proof to and replace $2$ by $p$.)

\textsc{Solution}

\input{q6.tex}


%==============================================================================
\newpage
Q7.
Let $p, q$ be distinct primes.
Prove that if $p \mid a$ and $q \mid a$, then $pq \mid a$.
(This was used in the proof of RSA in class.)

You must only use fact in the notes.
\begin{enumerate}[nosep]
\item Definition of divisibility
\item Basic properties of divisibility such as $\pm 1 | a$ for all $a$
and linearity of divisibility.
\item Euclidean property
\item Euclid's lemma
\item Extended Euclidean property
\item Fundamental theorem of arithmetic
\end{enumerate}

\textsc{Solution}

\input{q7.tex}


\end{document}
